### 2017-10-11 Sections/state_of_the_art.tex

Many quantum information experiments based on fibre networks \cite{sun2016quantum} have been performed using single-longitudinal-mode lasers, such as distributed feedback (DFB) lasers. The DFB laser has a stable wavelength that is etched by a grating, and can only be tuned slightly with temperature. Thus they are widely used in optical communication applications, such as dense wavelength division multiplexing (DWDM), where it is desired to have a tuneable signal and extremely narrow line width. Recently, several QKD schemes \cite{choi2011quantum, wang2015experimental} have been designed and implemented using multi-longitudinal-mode Fabry-Perot (FP) lasers, mainly because their costs are significantly lower than DFB lasers.

@@

Multi-photon GHZ states based on SPDC \cite{kwiat1995new} can be step-by-step constructed with two-photon entangled states. Here we take the four photon one as an example \cite{pan2012multiphoton}. Assume that two SPDC sources emit two polarisation entangled states \mbox{$\frac{1}{2}(\ket{HH} + \ket{VV})^{\otimes 2}$}. After passing through the polarizing beam-splitter (PBS), only the superposition \mbox{$\frac{1}{\sqrt{2}}(\ket{HHHH} + \ket{VVVV})$}, which is a four-photon GHZ state, leads to fourfold coincidence. In 1999, the world's first multi-particle 3-photon entanglement was generated \cite{bouwmeester1999observation, pan2000experimental}. After that, Pan \textit{et al.} broke the records continuously, and realised 4- \cite{zhao2003experimental}, 5- \cite{zhao2004experimental}, 6- \cite{lu2007experimental}, 8- \cite{yao2012observation}, 10- \cite{bib:tenPhotEnt} photon entanglement, and ten-qubit hyper-entanglement with two degrees of freedom \cite{gao2010experimental}.

Optical cluster states have been realised with 4- \cite{walther2005experimental}, 5- \cite{lu2008experimental}, 6- \cite{lu2007experimental}, 8- \cite{yao2012experimental} photon, and 2-photon 4-qubit \cite{chen2007experimental} and 6-qubit \cite{ceccarelli2009experimental}.

@@

Continuous-variable entangled states are attractive because they exploit standard optical modulation and measurement equipment \cite{ralph2009bright}. One mode squeezed vacuum state can be produced by a degenerate optical parametric amplifier. For cat states \mbox{$\propto(\ket{\alpha} - \ket{-\alpha})$} with small `size' \mbox{$|\alpha|^2 \lesssim 1$}, it can be generated by reflecting a small fraction of squeezed vacuum state toward an avalanche photodiode (APD) \cite{neergaard2006generation, ourjoumtsev2006generating, wakui2007photon}. The APD will herald the subtraction of one photon, thus project the transmitted beam into the desired state. For large `size' cat state (e.g., \mbox{$|\alpha|^2 > 2.3$}), it can be prepared by reflecting a half of photon-number state $\ket{n}$ toward a momentum quadrature homodyne detector \cite{ourjoumtsev2007generation,takahashi2008generation}. If the measurement outcome is close to 0, the transmitted mode is successfully projected into the desired state. Besides, two partite entangled cat states have been prepared and characterised \cite{ourjoumtsev2009preparation}. Initial demonstrations of the distillation of Gaussian entanglement have also been made \cite{takahashi2010entanglement, xiang2010heralded}.

@@

A variety of techniques have been developed to create squeezing and entanglement in atomic ensembles. The main methods exploit atom-light interactions in cold gases, or interaction between particles such as atom-atom collisions in Bose-Einstein condensates, or combined electrostatic and ion-light interaction in ion chains. Atom-light interaction currently represents the most mature method, and owns the highest squeezing of 20.1dB via an optical-cavity-based measurement \cite{hosten2016measurement}. Non-Gaussian states have also been produced recently in atomic ensembles. For instance, the detection of a single photon prepares almost 3000 atoms to an entangled Dicke state \cite{mcconnell2015entanglement}. Different atomic ensembles can be entangled. For example, two ensembles can be entangled by storage of two entangled light fields \cite{lukin2000entanglement}, four quantum memories can be entangled via spin wave \cite{choi2010entanglement}, large ensembles of up to $10^4$ pair-correlated atoms in Bose-Einstein condensates can be entangled to a twin-Fock state \cite{lucke2011twin}. More contents can be found in some excellent review papers \cite{kimble2008quantum, hammerer2010quantum, sangouard2011quantum, pezze2016non}.

@@

In quantum dots, the recombination of a pair of electrons and hole, named exciton, emits a single photon. There are also multiple electrons and/or holes leading to other transitions \cite{lodahl2015interfacing}. The simplest example is charged excitons, also called trions. Spontaneous emission from the trion state prepares a single electron spin in the quantum dot, while spin-photon entanglement was generated from such decay \cite{de2012quantum, gao2012observation}. Another example is a two electron-hole pairs, also called biexciton. Two photon polarisation-entangled states can be generated from biexciton-exciton cascade radiative decay in a single quantum dot \cite{muller2014demand}. The entangled photon pairs simultaneously show high purity of \mbox{$g_2(0) = 0.004$}, high fidelity of 0.81, high two-photon interference visibilities of 0.86 and on-demand generation efficiency of 0.86. Besides, a family of pyramidal site-controlled quantum dots allow areas with up to 15\% of polarisation-entangled photon emitters per chip to be obtained, with fidelities as high as 0.72 \cite{juska2013towards, mohan2010polarization}. Most recently, a cluster state of entangled five sequential photons was generated by periodic timed excitation of a matter qubit \cite{schwartz2016deterministic}. In each period, an entangled photon is added to the cluster state formed by the matter qubit and the previously emitted photons. More details about interfacing single photons and single quantum dots with photonic nanostructures can be found in several review papers \cite{de2013ultrafast, urbaszek2013nuclear, lodahl2015interfacing}.

@@

Because NV centres couple to both optical and microwave fields, they can also be used as a quantum interface between optical and solid-state systems. For example, quantum entanglement between a polarised optical photon and a NV center qubit has been realised in experiment \cite{togan2010quantum}.

@@

\subsubsection{Homodyning} \index{Homodyning}

@@

Alternatively, light can be stored in a high-Q cavity. The light cycles back and forth between the reflecting boundaries, allowing it to be injected into and retrieved from the cavity \cite{pittman2002single, pittman2002cyclical, leung2006quantum, maitre1997quantum, tanabe2007trapping, tanabe2009dynamic}. Unfortunately, the storage of light in cavities suffers from a tradeoff between short cycle time and long storage time, which limits the efficiency. Therefore, whereas optical delay lines and nanocavities could be appropriate for obtaining on-demand single photons from heralded sources \cite{saglamyurek2015quantum, jin2015telecom}, they may not be suitable for quantum memory or quantum repeaters.

@@

Experimentalists strived either to push further the quantumness of the memories, or to improve their figures of merit. On the way towards the greatest quantumness, EIT in cold atomic ensembles was used to store entangled photon pairs \cite{Choi2008mapping} as well as squeezed vacuum pulses \cite{appel2008quantum, honda2008storage}. More recently, it was used to store a quantum bit encoded in the polarization of a single photon with memory times hitting the millisecond range \cite{lettner2011remote, riedl2012bose, xu2013long}. On the way towards better and better figures of merit, recent experiments performed on classical signals have been reported with storage times up to the regime of one minute \cite{heinze2013stopped} or with efficiencies reaching 78\% \cite{chen2013coherent}. The (spatially) multimode nature of these storage media have also been probed experimentally \cite{ding2013single}.

There are also other promising approaches to quantum memories, e.g., quantum dots or crystalline defect centres. We note that the encoding of quantum information in the spin of individual nuclei in an otherwise spin-free or spin-polarised crystal lattice can be achieved with remarkably long coherence times \cite{steger2012quantum, saeedi2013room}, with a current record of several hours for the storage of strong light pulses in a spin ensemble \cite{zhong2015optically}. The major remaining challenge is to efficiently couple individual spins to photonic quantum channels. Typically, this requires exquisite control over the involved spatiotemporal light mode, posing a formidable challenge to system design and nanofabrication \cite{reiserer2015cavity}.

>>>>

Many quantum information experiments based on fibre networks \cite{sun2016quantum} have been performed using single-longitudinal-mode lasers, such as distributed feedback (DFB) lasers. The DFB laser has a stable wavelength that is etched by a grating, and can only be tuned slightly with temperature. Thus they are widely used in optical communication applications, such as dense wavelength division multiplexing (DWDM), where it is desired to have a tuneable signal and extremely narrow line width. Recently, several QKD schemes \cite{choi2011quantum, wang2015experimental} have been designed and implemented using multi-longitudinal-mode Fabry-Perot (FP) lasers\index{Fabry-Perot lasers}, mainly because their costs are significantly lower than DFB lasers.

@@

Multi-photon GHZ states based on SPDC \cite{kwiat1995new} can be step-by-step constructed with two-photon entangled states. Here we take the four photon one as an example \cite{pan2012multiphoton}. Assume that two SPDC sources emit two polarisation entangled states \mbox{$\frac{1}{2}(\ket{HH} + \ket{VV})^{\otimes 2}$}. After passing through the polarizing beam-splitter (PBS), only the superposition \mbox{$\frac{1}{\sqrt{2}}(\ket{HHHH} + \ket{VVVV})$}, which is a four-photon GHZ state, leads to fourfold coincidence. In 1999, the world's first multi-particle 3-photon entanglement was generated \cite{bouwmeester1999observation, pan2000experimental}. After that, Pan \textit{et al.} broke the records continuously, and realised 4- \cite{zhao2003experimental}, 5- \cite{zhao2004experimental}, 6- \cite{lu2007experimental}, 8- \cite{yao2012observation}, 10-photon entanglement \cite{bib:tenPhotEnt}, and 10-qubit hyper-entanglement with two degrees of freedom \cite{gao2010experimental}.

Optical cluster states have been realised with 4- \cite{walther2005experimental}, 5- \cite{lu2008experimental}, 6- \cite{lu2007experimental}, 8-photon \cite{yao2012experimental}, and 2-photon 4-qubit \cite{chen2007experimental} and 6-qubit \cite{ceccarelli2009experimental}.

@@

Continuous-variable entangled states are attractive because they exploit standard optical modulation and measurement equipment \cite{ralph2009bright}. One mode squeezed vacuum state can be produced by a degenerate optical parametric amplifier. For cat states \mbox{$\propto(\ket{\alpha} - \ket{-\alpha})$} with small `size' \mbox{$|\alpha|^2 \lesssim 1$}, it can be generated by reflecting a small fraction of squeezed vacuum state toward an avalanche photodiode (APD) \cite{neergaard2006generation, ourjoumtsev2006generating, wakui2007photon}. The APD will herald the subtraction of one photon, thus project the transmitted beam into the desired state. For large `size' cat state (e.g., \mbox{$|\alpha|^2 > 2.3$}), it can be prepared by reflecting a half of photon-number state $\ket{n}$ toward a momentum quadrature homodyne detector\index{Homodyne detection} \cite{ourjoumtsev2007generation,takahashi2008generation}. If the measurement outcome is close to 0, the transmitted mode is successfully projected into the desired state. Besides, two partite entangled cat states have been prepared and characterised \cite{ourjoumtsev2009preparation}. Initial demonstrations of the distillation of Gaussian entanglement have also been made \cite{takahashi2010entanglement, xiang2010heralded}.

@@

A variety of techniques have been developed to create squeezing and entanglement in atomic ensembles. The main methods exploit atom-light interactions in cold gases, or interaction between particles such as atom-atom collisions in Bose-Einstein condensates, or combined electrostatic and ion-light interaction in ion chains. Atom-light interaction currently represents the most mature method, and owns the highest squeezing of 20.1dB via an optical-cavity-based measurement \cite{hosten2016measurement}. Non-Gaussian states have also been produced recently in atomic ensembles. For instance, the detection of a single photon prepares almost 3000 atoms to an entangled Dicke state\index{Dicke state} \cite{mcconnell2015entanglement}. Different atomic ensembles can be entangled. For example, two ensembles can be entangled by storage of two entangled light fields \cite{lukin2000entanglement}, four quantum memories can be entangled via spin wave \cite{choi2010entanglement}, large ensembles of up to $10^4$ pair-correlated atoms in Bose-Einstein condensates can be entangled to a twin-Fock state \cite{lucke2011twin}. More contents can be found in some excellent review papers \cite{kimble2008quantum, hammerer2010quantum, sangouard2011quantum, pezze2016non}.

@@

In quantum dots, the recombination of a pair of electrons and hole, named exciton, emits a single photon. There are also multiple electrons and/or holes leading to other transitions \cite{lodahl2015interfacing}. The simplest example is charged excitons, also called trions. Spontaneous emission from the trion state prepares a single electron spin in the quantum dot, while spin-photon entanglement was generated from such decay \cite{de2012quantum, gao2012observation}. Another example is a two electron-hole pairs, also called biexciton. Two photon polarisation-entangled states can be generated from biexciton-exciton cascade radiative decay in a single quantum dot \cite{muller2014demand}. The entangled photon pairs simultaneously show high purity of \mbox{$g_2(0) = 0.004$}, high fidelity of 0.81, high two-photon interference visibilities of 0.86 and on-demand generation efficiency of 0.86. Besides, a family of pyramidal site-controlled quantum dots allow areas with up to 15\% of polarisation-entangled photon emitters per chip to be obtained, with fidelities as high as 0.72 \cite{juska2013towards, mohan2010polarization}. Most recently, a cluster state of entangled five sequential photons was generated by periodic timed excitation of a matter qubit \cite{schwartz2016deterministic}. In each period, an entangled photon is added to the cluster state formed by the matter qubit and the previously emitted photons. More details about interfacing single photons and single quantum dots with photonic nano-structures can be found in several review papers \cite{de2013ultrafast, urbaszek2013nuclear, lodahl2015interfacing}.

@@

Because NV centres couple to both optical and microwave fields, they can also be used as a quantum interface between optical and solid-state systems. For example, quantum entanglement between a polarised optical photon and a NV centre qubit has been realised in experiment \cite{togan2010quantum}.

@@

\subsubsection{Homodyning} \index{Homodyne detection}

@@

Alternatively, light can be stored in a high-Q cavity. The light cycles back and forth between the reflecting boundaries, allowing it to be injected into and retrieved from the cavity \cite{pittman2002single, pittman2002cyclical, leung2006quantum, maitre1997quantum, tanabe2007trapping, tanabe2009dynamic}. Unfortunately, the storage of light in cavities suffers from a tradeoff between short cycle time and long storage time, which limits the efficiency. Therefore, whereas optical delay lines and nano-cavities could be appropriate for obtaining on-demand single photons from heralded sources \cite{saglamyurek2015quantum, jin2015telecom}, they may not be suitable for quantum memory or quantum repeaters.

@@

Experimentalists strived either to push further the quantumness of the memories, or to improve their figures of merit. On the way towards the greatest quantumness, EIT in cold atomic ensembles was used to store entangled photon pairs \cite{Choi2008mapping} as well as squeezed vacuum pulses \cite{appel2008quantum, honda2008storage}. More recently, it was used to store a quantum bit encoded in the polarisation of a single photon with memory times hitting the millisecond range \cite{lettner2011remote, riedl2012bose, xu2013long}. On the way towards better and better figures of merit, recent experiments performed on classical signals have been reported with storage times up to the regime of one minute \cite{heinze2013stopped} or with efficiencies reaching 78\% \cite{chen2013coherent}. The (spatially) multimode nature of these storage media have also been probed experimentally \cite{ding2013single}.

There are also other promising approaches to quantum memories, e.g., quantum dots or crystalline defect centres. We note that the encoding of quantum information in the spin of individual nuclei in an otherwise spin-free or spin-polarised crystal lattice can be achieved with remarkably long coherence times \cite{steger2012quantum, saeedi2013room}, with a current record of several hours for the storage of strong light pulses in a spin ensemble \cite{zhong2015optically}. The major remaining challenge is to efficiently couple individual spins to photonic quantum channels. Typically, this requires exquisite control over the involved spatiotemporal light mode, posing a formidable challenge to system design and nano-fabrication \cite{reiserer2015cavity}.
